Real-Time Activity Classification by Match Filtering using Body-Worn Accelerometers

### 2016-06-27 notes/paper/paper.tex
Incremental changes.

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Results}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the algorithm performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user.
A separate instance of walking data has been used specifically for training the classifier.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced with decimation 8. Figures (c) and (d) are the corresponding ROC curves produced from (a) and (b) respectivley.}
  \label{fig:MF_performance}
\end{figure}
%
Figures \ref{fig:frac_walking_jogging} and \ref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\FloatBarrier
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instrument - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with a sufficiently large dynamic range that fixed-point processing cannot provide.
The 64 kB of SRAM provided sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which allows the microcontroller to have a high performance by allowing a high number of instruction cycles that ultimately allows for faster code execution.

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.45\textwidth]{32Hz_4sec_8downsample_2Reference_Pulse.PNG}
   \caption{Test setup}
   \label{fig:time_measure}
\end{figure}
%
The “Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer” oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-fitlering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows execution time over a $4.0$ second window at decimation $8$ while varying the sampling frequency and number of references. From this, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\%.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}

>>>>

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{Results}
%
\subsection{Algorithm Performance}
%
The Kinetisense sensor was used for the algorithm performance analysis in this paper.
This unit has a three-axis accelerometer with an acceleration range of -5g to +5g and a sampling rate of 128Hz.
Samples of walking, jogging, and bicycling have been recorded with this unit attached to the ankle of the user.
A separate instance of walking data has been used specifically for training the classifier.
%
\begin{figure}[!ht]
  %\captionsetup[subfigure]{labelformat=empty}
  \centering
  \subfloat[\label{fig:frac_walking_jogging}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_jogging.png}}
  \quad
  \subfloat[\label{fig:frac_walking_bicycling}]{\includegraphics[width=0.45\textwidth]{fraction_threshold_walking_to_bicycling.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_jogging}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_run.png}}
  \centering
  \quad
  \subfloat[\label{fig:roc_walking_bicycling}]{\includegraphics[width=.45\textwidth]{ROC_walk_to_bike.png}}
  \caption{Matched-filter performance comparing jogging and bicycling data against a reference for walking. Results produced at decimation 8. Figures (c) and (d) are the corresponding ROC curves produced from (a) and (b) respectivley.}
  \label{fig:MF_performance}
\end{figure}
%
Figures \ref{fig:frac_walking_jogging} and \ref{fig:frac_walking_bicycling} show the fraction of the correlation results that exceed the given correlation threshold when comparing the walking data reference to the given activity.
As we would expect, walking data matches better to walking data than the other activities.
ROC curves were produced from these corresponding results in \subref{fig:roc_walking_jogging} and \subref{fig:roc_walking_bicycling} to illustrate the effectiveness of the matched-filter as a classifier.
Figure \ref{fig:MF_performance} shows the comparison to bicycling.
Because the bicycling motions are considerably different from walking relative to jogging, we would expect better performance from bicycling.
%
%\FloatBarrier
%
\subsection{Throughput Performance}
%
To assess the real-time processing potential of the algorithms outlined in this paper, a throughput assessment was performed on the MSP432P401R microcontroller with varying parameters.
The Texas Instruments - MSP432P401R is a low power plus performance microcontroller that is based on the ARM 32-Bit Cortex-M4F processor.
Some of the notable features of the MSP432P410R is having a clock frequency of 48 MHz, 64 kB SRAM, 256 kB Flash Memory, and has an IEEE 754-compliant single-precision Floating-Point Unit (FPU) and Memory Protection Unit (MPU).
The single-precision processing capability on this unit provides us with a sufficiently large dynamic range necessary to process the algorithms that fixed-point processing cannot provide.
The 64 kB of SRAM provides sufficient space to process the results found in figure \ref{fig:timing} and
the 256 kB is more than enough flash memory to store the program and data references.
The clock frequency operates at 48 MHz which determines the execution speed.

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.45\textwidth]{32Hz_4sec_8downsample_2Reference_Pulse.PNG}
   \caption{Oscilloscope output used for measuring the processing time of the algorithms on the MSP432. This is a case setup of 32Hz sampling frequency over a 4 second time window with 2 references at decimation 8 showing the voltage output over time. Execution time is measured from the start to end of the ``high'' (3.3 Volts) time interval.}
   \label{fig:time_measure}
\end{figure}
%
The ``Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer'' oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-filtering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.
An example can be seen in Figure \ref{fig:time_measure}.

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds.}
   \label{fig:timing}
\end{figure}
%
Randomized data was used for the purpose of analysing the throughput performance on the device while varying the parameters.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process while operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even powered factors, one can observe better throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even valued time lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
conclusion text.
%
\appendices
%
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi
%
\bibliographystyle{plain}
\bibliography{reference}
\end{document}
